### 2013-07-23 Tex/chapters/methods.tex
Korrekturen betreffend H (\hat H)

\subsection{Thermodynamic Theory}
\label{ssec:theory}
    %~ To examine the disordered Ising model in Equilibrium,
    %~ Its temperature \(T\),
    %~ volume and number of particles \(N\) are constant, so that it is a canonical
    %~ ensemble.
    In a canonical ensemble the probability \(p_i\) of a state
    \(i\) is distributed according to a Boltzman distribution
    \[p_i \propto e^{-\frac{\hat{H}}{k_B T}}\]
    Further the free energy \(F\) of an canonical
    ensemble is minimized and all observables can be derived from \(F\)
    in a straight forward way, as stated in every textbook about
    statistical physics (e.g. \cite{nolting2005}).\\
    Because of \(F=U-TS\) where \(U=\avg{H}\) is the internal energy and
    \(S\) entropy, one can guess, that for low \(T\) the internal energy
    will be low, and for big \(T\) the entropy will be high, to minimize
    \(F\). The Hamiltonian of the Ising model creates an conflict between
    \(U\) and \(S\) by assigning low energies to states with high order.
    So \(U\) is low if all spins are aligned, where \(S\) as a measure of
    disorder is low. This preliminary considerations make a phase
    transition at some \(T\), where a high entropy lowers \(F\) more than
    a low internal energy, very plausible.\\\
    To determine
    \[F=-k_{B}T \ln{Z}\]
    one has to know the partition function
    \[Z=\sum_i e^{-\frac{\hat{H}_i}{k_{B}T}}\]
    where the sum goes over all possible states \(i\) of the system.
    For small \(N\) the partition function is computable, but the system
    may show very different properties than in the thermodynamic limit.
    To minimize these finite size effects, it is desirable to examine
    systems with large \(N\). But then there are of course next to
    infinity many different states, such that it is unfeasible to
    calculate the energy for each state, except for cases, where it is
    possible to solve it analytically.\\
    If not, one can get estimates of the observables for big \(N\) using
    Monte Carlo simulations.\\

@@

    \subsubsection{Metropolis}
        A Metropolis Monte Carlo\cite{Metropolis1953} simulation of an
        Ising model will choose a random spin, calculate the energy change
        \(\Delta \hat H\) defined in eq. \eqref{eq:dH} that would result
        from a flip of that spin and execute the flip with the probability \(A\)
        eq. \eqref{eq:metropolis} \cite{NewmanBarkema1999} \cite{Katzgraber2011}.
        \begin{equation}
            \Delta \hat H = \hat H(\nu) - \hat H(\mu)\\
            \label{eq:dH}
        \end{equation}
        \begin{equation}
            A(\mu \to \nu)
            \begin{cases}
                1                                 & \Delta \hat H \le 0 \\
                \exp{\brac{-\beta \Delta \hat H}} & \Delta \hat H > 0
            \end{cases}
            \label{eq:metropolis}
        \end{equation}
        So if a transition lowers the energy it is always done. This
        results in a high ratio between chosen spins and flipped spins.
        Therefore it minimizes the calculations needed for a change of
        the state.

>>>>

\subsection{Thermodynamic Theory}
\label{ssec:theory}
    %~ To examine the disordered Ising model in Equilibrium,
    %~ Its temperature \(T\),
    %~ volume and number of particles \(N\) are constant, so that it is a canonical
    %~ ensemble.
    In a canonical ensemble the probability \(p_i\) of a state
    \(i\) is distributed according to a Boltzman distribution
    \[p_i \propto e^{-\frac{H_i}{k_B T}}\]
    Further the free energy \(F\) of an canonical
    ensemble is minimized and all observables can be derived from \(F\)
    in a straight forward way, as stated in every textbook about
    statistical physics (e.g. \cite{nolting2005}).\\
    Because of \(F=U-TS\) where \(U=\avg{H}\) is the internal energy and
    \(S\) entropy, one can guess, that for low \(T\) the internal energy
    will be low, and for big \(T\) the entropy will be high, to minimize
    \(F\). The Hamiltonian of the Ising model creates an conflict between
    \(U\) and \(S\) by assigning low energies to states with high order.
    So \(U\) is low if all spins are aligned, where \(S\) as a measure of
    disorder is low. This preliminary considerations make a phase
    transition at some \(T\), where a high entropy lowers \(F\) more than
    a low internal energy, very plausible.\\\
    To determine
    \[F=-k_{B}T \ln{Z}\]
    one has to know the partition function
    \[Z=\sum_i e^{-\frac{H_i}{k_{B}T}}\]
    where the sum goes over all possible states \(i\) of the system.
    For small \(N\) the partition function is computable, but the system
    may show very different properties than in the thermodynamic limit.
    To minimize these finite size effects, it is desirable to examine
    systems with large \(N\). But then there are of course next to
    infinity many different states, such that it is unfeasible to
    calculate the energy for each state, except for cases, where it is
    possible to solve it analytically.\\
    If not, one can get estimates of the observables for big \(N\) using
    Monte Carlo simulations.\\

@@

    \subsubsection{Metropolis}
        A Metropolis Monte Carlo\cite{Metropolis1953} simulation of an
        Ising model will choose a random spin, calculate the energy change
        \(\Delta H\) defined in eq. \eqref{eq:dH} that would result
        from a flip of that spin and execute the flip with the probability \(A\)
        eq. \eqref{eq:metropolis} \cite{NewmanBarkema1999} \cite{Katzgraber2011}.
        \begin{equation}
            \Delta H = H(\nu) - H(\mu)\\
            \label{eq:dH}
        \end{equation}
        \begin{equation}
            A(\mu \to \nu)
            \begin{cases}
                1                            & \Delta H \le 0 \\
                \exp{\brac{-\beta \Delta H}} & \Delta H > 0
            \end{cases}
            \label{eq:metropolis}
        \end{equation}
        So if a transition lowers the energy it is always done. This
        results in a high ratio between chosen spins and flipped spins.
        Therefore it minimizes the calculations needed for a change of
        the state.

### 2013-07-23 Tex/chapters/model.tex
Korrekturen betreffend H (\hat H)

\subsection{The Disordered Ising Model}
\label{ssec:isingmodel}
    The examined model is a modified 2D Ising model.
    First, let us define the Ising model and then explain the modifications.
    The Ising model is a square lattice with edge length \(L\) and
    \(N=L^2\) sites. Each site has a magnetic moment, the spin. Each
    spin can take a value \(s \in \{-1,+1\}\) and interacts with its
    nearest neighbors described by the Hamiltonian from eq. \eqref{eq:hamiltonian}
    \begin{equation}
        H = \sum_{\avg{i,j}}J_{ij}s_{i}s_{j} - \tilde{H} \sum_i s_i
        \label{eq:hamiltonian}
    \end{equation}
    \(\avg{i,j}\) refers to nodes \(i\) and \(j\) which are nearest
    neighbors. And \(J_{ij}\) is the coupling constant between \(i\) and
    \(j\). If \(J_{ij} > 0 \forall i,j\) the model resembles a ferromagnet.
    \(\tilde{H}\) denotes the outer magnetic field.\\
    In this thesis \(\tilde{H}=0\). The most important modification is, that
    the sites of the square lattice are displaced. The
    displacement is randomly Gauß distributed with the standard
    deviation \(\sigma\), i.e. the \(x\) and \(y\) coordinates of the
    sites are displaced by random \(\Delta x\) and \(\Delta y\) drawn
    from a Gauß distribution eq. \eqref{eq:gauss}.
    This is sketched in fig. \ref{fig:displacement}.
    \begin{equation}
        f(x)=\frac{1}{\sqrt{2\pi}\sigma}\mathrm{e}^{-\frac{x^2}{2\sigma^2}}
        \label{eq:gauss}
    \end{equation}
    \[x \to x + \Delta x\]
    \[y \to y + \Delta y\]
    \begin{figure}[htbp]
        \centering
        \input{images/Displacement}
        \caption[Sketch how the Displacement Works]
        {
            Sketch how the displacement of the nodes works. The nodes
            get displaced by \(\Delta x\) and \(\Delta y\) drawn from the
            distributions displayed next to the points. The original
            square lattice is indicated by dashed lines.
        }
        \label{fig:displacement}
    \end{figure}\\
    This \(\sigma\) is also called \emph{disorder parameter} in the following.
    Because most sites will only have one nearest neighbor after the
    displacement, the lattice would collapse to many very small clusters.
    To avoid this, the new "nearest" neighbors are those sites connected
    by an edge. The edges are constructed according to
    one of the two in section \ref{ssec:graphtypes} defined rules,
    so that the lattice represents a proximity graph. Note that edges
    of a proximity graph do not cross each other, hence the 2D character
    of the lattice is preserved. The coupling constant \(J\) gets
    identified with edge weights. The weight of an edge \(E_{ij}\) is
    \(J_{ij} = \exp (\alpha (1-d_{ij}))\) where \(d_{ij}\) is the Euclidean
    distance between the nodes \(i\) and \(j\). The free parameter
    \(\alpha\) is set to \(\alpha = 0.5\) inspired by \cite{Lima2000}.
    The boundary is periodic e.g. nodes near the right edge can be
    connected to nodes near the left edge and vice versa. Analogous the
    top and bottom edges are connected. One can imagine that the model
    lives on the surface of a torus.\\
    For \(\sigma = 0\) this is the standard Ising model with \(J = 1\),
    for which exists an analytic solution \cite{Onsager1944}. And for
    \(\sigma \gtrsim 1\) the nodes are distributed randomly. This case
    is studied for constant \(J\) on the Delaunay triangulation\cite{Janke1994}.\\

>>>>

\subsection{The Disordered Ising Model}
\label{ssec:isingmodel}
    The examined model is a modified 2D Ising model.
    First, let us define the Ising model and then explain the modifications.
    The Ising model is a square lattice with edge length \(L\) and
    \(N=L^2\) sites. Each site has a magnetic moment, the spin. Each
    spin can take a value \(s \in \{-1,+1\}\) and interacts with its
    nearest neighbors described by the Hamiltonian from eq. \eqref{eq:hamiltonian}
    \begin{equation}
        H = \sum_{\avg{i,j}}J_{ij}s_{i}s_{j} - \sum_i \tilde{H}_i s_i
        \label{eq:hamiltonian}
    \end{equation}
    \(\avg{i,j}\) refers to nodes \(i\) and \(j\) which are nearest
    neighbors. And \(J_{ij}\) is the coupling constant between \(i\) and
    \(j\). If \(J_{ij} > 0 \ \forall i,j\) the model resembles a ferromagnet.
    \(\tilde{H}_i\) denotes the outer magnetic field at the position of
    site \(i\).\\
    In this thesis \(\tilde{H}_i=0 \ \forall i\). The most important
    modification is, that the sites of the square lattice are displaced.
    The displacement is randomly Gauß distributed with the standard
    deviation \(\sigma\), i.e. the \(x\) and \(y\) coordinates of the
    sites are displaced by random \(\Delta x\) and \(\Delta y\) drawn
    from a Gauß distribution eq. \eqref{eq:gauss}.
    This is sketched in fig. \ref{fig:displacement}.
    \begin{equation}
        f(x)=\frac{1}{\sqrt{2\pi}\sigma}\mathrm{e}^{-\frac{x^2}{2\sigma^2}}
        \label{eq:gauss}
    \end{equation}
    \[x \to x + \Delta x\]
    \[y \to y + \Delta y\]
    \begin{figure}[htbp]
        \centering
        \input{images/Displacement}
        \caption[Sketch how the Displacement Works]
        {
            Sketch how the displacement of the nodes works. The nodes
            get displaced by \(\Delta x\) and \(\Delta y\) drawn from the
            distributions displayed next to the points. The original
            square lattice is indicated by dashed lines.
        }
        \label{fig:displacement}
    \end{figure}\\
    This \(\sigma\) is also called \emph{disorder parameter} in the following.
    Because most sites will only have one nearest neighbor after the
    displacement, the lattice would collapse to many very small clusters.
    To avoid this, the new "nearest" neighbors are those sites connected
    by an edge. The edges are constructed according to
    one of the two in section \ref{ssec:graphtypes} defined rules,
    so that the lattice represents a proximity graph. Note that edges
    of a proximity graph do not cross each other, hence the 2D character
    of the lattice is preserved. The coupling constant \(J\) gets
    identified with edge weights. The weight of an edge \(E_{ij}\) is
    \(J_{ij} = \exp (\alpha (1-d_{ij}))\) where \(d_{ij}\) is the Euclidean
    distance between the nodes \(i\) and \(j\). The free parameter
    \(\alpha\) is set to \(\alpha = 0.5\) inspired by \cite{Lima2000}.
    The boundary is periodic e.g. nodes near the right edge can be
    connected to nodes near the left edge and vice versa. Analogous the
    top and bottom edges are connected. One can imagine that the model
    lives on the surface of a torus.\\
    For \(\sigma = 0\) this is the standard Ising model with \(J = 1\),
    for which exists an analytic solution \cite{Onsager1944}. And for
    \(\sigma \gtrsim 1\) the nodes are distributed randomly. This case
    is studied for constant \(J\) on the Delaunay triangulation\cite{Janke1994}.\\
